### 2016-05-25 presentation.tex
Add picture of memory address space

\documentclass{beamer}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{filecontents}
\usepackage{mathtools}
\usepackage{underscore}
\usepackage{tikz}

\usetikzlibrary{patterns,shapes,positioning,calc}
\setbeamertemplate{bibliography item}[text]
\lstset{language=C, basicstyle=\ttfamily\scriptsize, columns=flexible, belowskip=0pt}
\newcommand{\perfctr}[1] {
  {\lowercase{#1}}
}

\title{Measurement Bias from Address Aliasing}

\author[Lars Kirkholt Melhus, Rune Erlend Jensen]
{Lars Kirkholt Melhus \and Rune Erlend Jensen}
\institute{
  Dept. of Computer and Information Science\\
  Norwegian University of Science and Technology \\
  Trondheim, Norway
}

\date[2016] % (optional)
{International Workshop on Automatic Performance Tuning, Chicago, May 2016}

\subject{Computer Science}

% Allocated to each presentation: 30 mins. Probably 20 mins + questions,
% a couple of minutes per slide should mean around 10 slides at the most.

\begin{document}

\frame{\titlepage}

\begin{frame}

\frametitle{Outline}

\begin{itemize}
  \item Identify address aliasing as an underlying explanation for measurement bias.
  \item See how aliasing affects dynamic memory allocation.
  \item Look at opportunities to optimize for aliasing.
  \item Up to 2x speedup.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Measurement Bias}

% Show example code and plot from Mytkowitz.

\begin{itemize}
  \item Runtime performance change from external factors.
  \begin{itemize}
    \item Environment variables
    \item Link ordering
  \end{itemize}
  \item Challenge for performance analysis.
  \item Effects can invalidate perceived speedup~\cite{Mytkowicz:2008:OE&MB}.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Methodology}

\begin{columns}[T] % align columns
\begin{column}{.40\textwidth}

  % Some points on hardware perf counters.
  \begin{itemize}
    \item Control execution environment (disable ASLR).
    \item Instrument program using hardware performance counters.
    \item Measure over range of environments.
  \end{itemize}

\end{column}
\begin{column}{.60\textwidth}

\begin{figure}
  \begin{tikzpicture}[font=\footnotesize]
    % See page 453 in pgf manual
    \node [
      rectangle split, rectangle split parts=10, 
      rectangle split part fill={white, white, lightgray, white, lightgray, white, lightgray, white},
      draw, anchor=center, text width=1.8cm
    ] (m) {
        \nodepart{one}
          env. vars
        \nodepart{two}
          stack
        \nodepart{four}
          mmap area
        \nodepart{six}
          heap
        \nodepart{eight}
          bss
        \nodepart{nine}
          data
        \nodepart{ten}
          text
      };
    \draw [decorate, decoration={brace, amplitude=5pt}] (m.south west) -- (m.seven split west) 
      node [black, midway, xshift=-1.5cm, text width=2.0cm] 
        { Program code and static data } ;
    \node[right] at (m.north east)
      { \scriptsize{0x7fff'ffffffff} } ;
    \node[right] at (m.south east)
      { \scriptsize{0x400000} } ;
  \end{tikzpicture}
  \caption{Memory execution context, assuming a 64-bit process running on a Linux system.}
  \label{fig:virtualmemory}
\end{figure}

\end{column}
\end{columns}

\end{frame}

\begin{frame}[fragile, allowframebreaks]
\frametitle{Bias from Environment Size}

\begin{figure}
  \begin{lstlisting}[frame=single, xleftmargin=.01\textwidth, xrightmargin=.01\textwidth, aboveskip=-2pt]
static int i, j, k;

int main() {
    int g = 0, inc = 1;
    for (; g < 65536; g++) {
        i += inc;
        j += inc;
        k += inc;
    }
    return 0;
}
  \end{lstlisting}
  \caption{\label{fig:microkernel}Microkernel first presented by Mytkowicz \emph{et al.}\cite{Mytkowicz:2009:WrongData}, showing bias to environment size.}
\end{figure}

% Also an optimization problem (relevant for auto-tuning).

Execute program under different environments, setting dummy variable to $K$ zero characters.

\begin{lstlisting}[breaklines=true, belowskip=-10pt, basicstyle=\ttfamily\small]
$ env -i X=`head -c K </dev/zero | tr '\0' '0'` ./loop
\end{lstlisting}

\pgfplotstableread{bin/microkernel-cycles-haswell.dat}{\stackoffsettable}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}[
        title = Haswell,
        width = \textwidth,
        height = 4.7cm,
        font = \footnotesize,
        xlabel=Bytes added to environment,
        ylabel=Cycles,
        domain = 0:8192,
        xtick = {0,1024,...,8192},
        xmin = 0,
        xmax = 8192,
        cycle list name = exotic
      ]
      \addplot[ycomb] table[x expr = \thisrowno{0}*16, y = cycles:u] \stackoffsettable ;
    \end{axis}
  \end{tikzpicture}
  \caption{Measured cycle count samples for 512 different environments.}
\end{figure}

% Also an optimization problem (relevant for auto-tuning).

\begin{table}
  %\centering
  \caption{Events with significant correlation to cycle count.\label{tab:loopcorrelation}}
  \pgfplotstabletypeset[
    int detect, % Output whole numbers for counter values
    col sep=comma,
    columns={Performance counter, Median, [index]3, [index]4},
    column type=r,
    font=\footnotesize,
    columns/Performance counter/.style={
      string type, 
      column type=l,
      column type/.add={|}{},
      postproc cell content/.append code={
        \pgfkeysalso{@cell content=\perfctr{##1}}
      }
    },
    every head row/.style={
      output empty row,
      before row={\hline
        Performance counter & Median & Spike 1 & Spike 2 \\
      },
      after row=\hline\hline
    },
    every last row/.style={after row=\hline},
    every last column/.style={column type/.add={}{|}}
  ]{bin/microkernel-comparison-haswell.csv}
\end{table}

\end{frame}

\begin{frame}
\frametitle{4K Address Aliasing}

\begin{columns}[T] % align columns
\begin{column}{.50\textwidth}

\begin{itemize}
  %\item Collisions between static and automatic variables, depending on alignment of stack.
  \item Addresses of static i, j and k are fixed for all environments.
  \item Addresses of g and inc are pushed down with size of environment variables.
  \item Spike when address of inc is \emph{aliasing} with i in the last 12 bits.
  \item Performance impact from \emph{false dependencies} in CPU out of order execution.
\end{itemize}

\end{column}
\hfill
\begin{column}{.50\textwidth}

% Virtual memory/randomization illustration
\begin{figure}
  \begin{tikzpicture}[font=\footnotesize]
    % See page 453 in pgf manual
    \node [
      rectangle split, rectangle split parts=5, 
      rectangle split part fill={white, white, lightgray, white, lightgray},
      draw, anchor=center, text width=2.5cm
    ] (m) {
      \nodepart{one}
        \scriptsize{USER=lars;SHELL
        =/bin/bash;X=0000
        00000000000000...}
      \nodepart{two}
        .stack
        \begin{flushright}
        inc \\
        g
        \end{flushright}
      \nodepart{four}
        .bss
        \begin{flushright}
        k \\
        j \\
        i
        \end{flushright}
    };

    \node[right, yshift=0cm] at (m.two east) {
      \scriptsize{0x7fff'ffffe{\color{red}03c}} % inc
    };
    \node[right, yshift=-0.4cm] at (m.two east) {
      \scriptsize{0x7fff'ffffe038} % g
    };

    \node[right, yshift=0cm] at (m.four east) {
      \scriptsize{0x601044} % k
    };
    \node[right, yshift=-0.4cm] at (m.four east) {
      \scriptsize{0x601040} % j
    };
    \node[right, yshift=-0.8cm] at (m.four east) {
      \scriptsize{0x601{\color{red}03c}} % i
    };

    %\draw [decorate, decoration={brace, amplitude=5pt}] (m.south west) -- (m.seven split west) 
    %  node [black, midway, xshift=-1.5cm, text width=2.0cm] 
    %    { Program code and static data } ;

    %\node[right] at (m.north east) { \scriptsize{0x7fff'ffffffff} } ;
    %\node[right] at (m.south east) { \scriptsize{0x400000} } ;
  \end{tikzpicture}
  \caption{Addresses of each variable in the first spike.}
\end{figure}

% Note: Could have gotten more collisions if g also was colliding with something,
%       but not noticeable effect on cycles executed.

\end{column}
\end{columns}

\end{frame}

\begin{frame}[fragile]
  \frametitle{How to Measure Address Aliasing}

  \begin{description}
    \item[{\small LD\_BLOCKS\_PARTIAL.ADDRESS\_ALIAS.}] 
    \emph{``Counts the number of loads that have partial address match with preceding stores, causing the load to be reissued.''} 
    \cite[B.3.4.4]{OptimizationManual}
  \end{description}
  %This event is listed in the manual for microarchitectures going back to ``Nehalem'', including ``Ivy Bridge'' and ``Haswell'' CPUs~\cite{Volume3B}.
  %Older architectures such as Core do not have this particular event listed, but 4K aliasing is covered by a more general event:

  Same or similar counters found in architectures dating back through Core 2.

  % \begin{itemize}
  %   \item Older architectures, such as Core, have alias covered in a more general event:
  % \end{itemize}

  % \begin{description}
  %   \item[{\small LOAD\_BLOCKS.OVERLAP\_STORE.}] 
  %   \emph{``Loads that partially overlap an earlier store, or 4-Kbyte aliased with a previous store.''} 
  %   \cite[Table 19-17]{Volume3B}
  % \end{description}

  % Describe LD_BLOCKS_PARTIAL.ADDRESS_ALIAS
\end{frame}

% After this, done with method for detecting alias, and answers question
% from old paper. Now lets see what else can be discovered.

\begin{frame}[fragile]
\frametitle{Bias in Heap Allocation}

\begin{itemize}
  \item Alias conflicts can happen between any sections of memory.
  \item While stack addresses depend on environment, heap addresses depend on allocators.
\end{itemize}

\begin{table}
  \caption{Addresses returned by different heap allocators when allocating pairs of equally sized buffers. Same 12 bit suffix indicate an aliasing pair.}
  \pgfplotstabletypeset[
    font=\footnotesize,
    col sep=comma,
    string type,
    columns={Allocation, 5120, 1048576},
    column type=r,
    columns/Allocation/.style={
      string type, 
      column type=l,
      column type/.add={|}{}
    },
    every head row/.style={
      output empty row,
      before row={\hline
         & 5,120 B & 1,048,576 B \\
        },
      after row=\hline\hline,
    },
    every last row/.style={after row=\hline},
    every last column/.style={column type/.add={}{|}},
    every odd row/.style={after row=\hline},
  ]{bin/malloc-comparison.csv}
\end{table}

  % Motivation for looking: Determines addresses.

  % Short survay, most use mmap for large allocations.

  % punchline: mmap always alias.

\end{frame}

\begin{frame}[fragile]
  \frametitle{mmap}

Most allocators tend to use anonymous memory mapping for large requests.

\begin{lstlisting}[frame=single, xleftmargin=.01\textwidth, xrightmargin=.01\textwidth, belowskip=1cm]
$ man mmap

(...) on Linux, the mapping will be created at a nearby page boundary.
The address of the new mapping is returned as the result of the call.

\end{lstlisting}

Since page size is 4 KiB, such allocations will \emph{always} alias with each other.

\end{frame}

\begin{frame}[fragile, allowframebreaks]
  \frametitle{Optimization Resistant Code}

% Image of sliding window with buffers in and out (Color variables?)

Example program which is sensitive to aliasing between buffers, continuously generating false dependencies:

\begin{tikzpicture}[
  barstyle/.style={rectangle,fill=lightgray,draw,anchor=center,minimum width=9cm}
]
\node [barstyle] (in) at (0,1) { } ;
\node [barstyle] (out) at (0,0) { } ;
\node [rectangle,draw,minimum height=1.8cm] (bar) at (-2,0.5) { } ;

\node[] (input label) [left=of in] { in } ;
\node[] (output label) [left=of out] { out } ;

\draw[->] (-2,0.5) -- (-1.5,0.5) node [right,align=center] {\scriptsize{Sliding window}} ;

\end{tikzpicture}

\begin{figure}[t]
% Convolution example code
  \lstinputlisting[
    language=C,
    frame=single,
    xleftmargin=.00\textwidth,
    xrightmargin=.00\textwidth,
    aboveskip=-5pt
  ]{bin/convolution-kernel.c}
  \caption{Simplified implementation of convolution with a fixed kernel.}
  \label{lst:conv}
\end{figure}

% \begin{itemize}
%   \item Most heap allocators will \emph{always} align arrays on the same 12 bit suffix.
%   \item Manually adjusting addresses necessary to get good performance.
% \end{itemize}

Adjust address of output array by \texttt{d} bytes manually to avoid alising:

%mmap(NULL, (n + d), PROT_READ | PROT_WRITE, MAP_PRIVATE | MAP_ANONYMOUS, -1, 0) + d;
\begin{lstlisting}[belowskip=-4pt, basicstyle=\ttfamily\small]
out = malloc((N + d) * sizeof(float));
conv(N, in, out + d);
\end{lstlisting}

% The most awesome graph, O2 of convolution kernel
\pgfplotstableread{bin/conv-default-o2-haswell.estimate.dat}{\convtabletwohaswell}
\pgfplotstableread{bin/conv-default-o3-haswell.estimate.dat}{\convtablethreehaswell}
\begin{figure}[t]
  \begin{tikzpicture}
    \begin{axis}[
        title=Haswell -O2,
        font=\footnotesize,
        cycle list name=exotic,
        width=\textwidth/1.8,
        height=4.7cm,
        skip coords between index={18}{32} % Limit level of detail to fit page width nicely
      ]
      \addplot table[x expr = \thisrowno{0}, y = cycles:u] \convtabletwohaswell ;
      \addplot table[x expr = \thisrowno{0}, y = r0107:u ] \convtabletwohaswell ;
      \addlegendentry{Cycles} ;
      \addlegendentry{Alias} ;
    \end{axis}
  \end{tikzpicture}
  \begin{tikzpicture}
    \begin{axis}[
        title=Haswell -O3,
        font=\footnotesize,
        cycle list name=exotic,
        width=\textwidth/1.8,
        height=4.7cm,
        skip coords between index={18}{32}
      ]
      \addplot table[x expr = \thisrowno{0}, y = cycles:u] \convtablethreehaswell ;
      \addplot table[x expr = \thisrowno{0}, y = r0107:u ] \convtablethreehaswell ;
      \addlegendentry{Cycles} ;
      \addlegendentry{Alias} ;
    \end{axis}
  \end{tikzpicture}
  \caption{\label{fig:conv-default}Cycle- and alias counts for different offsets between input and output arrays in convolution kernel, input size $N=2^{20}$.}
\end{figure}

\end{frame}

\begin{frame}[fragile]
\frametitle{Aliasing in BLAS libraries}

% Maybe mention FFTW results from Master's?

Measured matrix-vector multiplication (cblas\_dgemv) with increasing offset between buffer addresses.
Computing $\boldsymbol{y}=A\boldsymbol{x}$.

\pgfplotstableread{bin/libblas.dat}{\libblastable}
\pgfplotstableread{bin/libatlas.dat}{\libatlastable}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}[
        title=libblas,
        height=4.7cm,
        font=\footnotesize,
        xlabel=Address offset of $\boldsymbol{y}$,
        %ylabel=Event count,
        cycle list name=exotic,
        width=\textwidth/1.8, % Make it fit to text width side by side
        legend style={at={(0.4,0.55)},anchor=west,draw=none}
      ]
      \addplot table[x expr = \thisrowno{0}, y = cycles:u] \libblastable ;
      \addplot table[x expr = \thisrowno{0}, y = r0107:u ] \libblastable ;
      \addlegendentry{Cycles} ;
      \addlegendentry{Alias} ;
    \end{axis}
  \end{tikzpicture}
  \begin{tikzpicture}
    \begin{axis}[
        title=libatlas,
        height=4.7cm,
        font=\footnotesize,
        xlabel=Address offset of $\boldsymbol{y}$,
        %ylabel=Event count,
        cycle list name=exotic,
        width=\textwidth/1.8,
        legend style={at={(0.4,0.55)},anchor=west,draw=none}
      ]
      \addplot table[x expr = \thisrowno{0}, y = cycles:u] \libatlastable ;
      \addplot table[x expr = \thisrowno{0}, y = r0107:u ] \libatlastable ;
      \addlegendentry{Cycles} ;
      \addlegendentry{Alias} ;
    \end{axis}
  \end{tikzpicture}
  \caption{\label{fig:blas}Cycles executed and address alias events for varying relative address offset between matrix $A$ and vector $\boldsymbol{y}$.}
  % Each sample point represents an increment of 16 bytes, or 0x10.}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Optimizing for address aliasing}

% Effects of address aliasing are real, these are some ideas for optimization:

\begin{itemize}
  \item Mark buffers with \texttt{restrict}.
  \begin{itemize}
    \item Can allow the optimizer to reduce number of memory accesses.
  \end{itemize}
  \item Use a special purpose allocator.
  \begin{itemize}
    \item Heuristics for avoiding pairwise aliasing buffers.
    \item No known implementation with this goal in mind, to our knowledge.
  \end{itemize}
  \item Explicitly adjust address offset.
  \begin{itemize}
    \item Consider memory address alignment for performance tuning inner loops.
    \item Buffer addresses as candidate for automatic performance tuning
  \end{itemize}
\end{itemize}

\end{frame}

\begin{frame}

\frametitle{References}

\scriptsize{\bibliographystyle{IEEEtran}}
\bibliography{references}

\end{frame}

\end{document}